---
title: "SRT Phenotype"
output:
  html_document:
    toc: TRUE
    depth: 3
editor_options: 
  chunk_output_type: console
chunk_output_type: console
---
***

## Read in dataset

```{r load_data}
data.geno.pheno <- read.csv("./010521_UKBB_HI_genotypeqc_expandedwhite_396974indiv_excr.csv") #cleaned data



library(dplyr)

```



### hearing aids (3393)
### step 1: classify cases and controls
```{r get cases and controls, warning=FALSE, collapse=FALSE, echo=T}
library(tidyverse)

hearing_all = data.geno.pheno %>% 
  select(IID,FID,f.31.0.0,f.34.0.0, f.21003.0.0,f.21003.1.0,f.21003.2.0,f.21003.3.0,f.3393.0.0,f.3393.1.0,f.3393.2.0,f.3393.3.0,f.2247.0.0,f.2247.1.0,f.2247.2.0,f.2247.3.0,f.2257.0.0, f.2257.1.0, f.2257.2.0, f.2257.3.0,starts_with("f.41270"),starts_with("f.41280")) 


#classify the cases and controls hearing aid, based on Z974
hearing_all = hearing_all %>% 
  mutate(cases_Z974 = apply(select(hearing_all,starts_with("f.41270")), 1, function(x) any(x %in% c("Z974"))
  )) 


#classify the cases and controls for hearing aid, based on 3393
hearing_all = hearing_all %>% 
  mutate(cases_3393 = apply(select(hearing_all,starts_with("f.3393")), 1, function(x) length(which(x == "Yes")) > 0 & max(which(x != "Yes")) < min(which(x == "Yes")))
  )


hearing_all = hearing_all %>% 
      mutate(control_3393 = ifelse(apply(hearing_all, 1, function(x){
              col_names = names(x)[grep("f.3393", names(x))]
              condition1 = sapply(col_names, function(y){
                x[y] %in% c("No",NA)
              })
              
              condition2 = sapply(col_names, function(y){
                x[y] %in% c(NA)
              })
              
              all(condition1) & !all(condition2)
            }),
            "FALSE",
            "NA"
            )
         ) 

                                     
#hearing_all$cases = apply(select(hearing_all,starts_with("f.3393")), 1, function(x) length(which(x == "Yes")) > 0 & max(which(x != "Yes")) < min(which(x == "Yes")))

#148 individuals are true for Z974 but controls for 3393
hearing_aid_trial <- hearing_all %>% 
  filter(cases_Z974 == "TRUE" & control_3393 == "FALSE")

hearing_all <- hearing_all %>% 
  mutate(reclass_aid= case_when(
    cases_Z974 == "TRUE" & control_3393 == "FALSE" ~ "TRUE",
    cases_Z974 == "TRUE" & control_3393 == "NA"  ~ "TRUE",
    TRUE ~"FALSE") ) 

#13808 individuals are cases for either Z947 or 3393
hearing_aid_cases <- hearing_all %>% 
 filter(reclass_aid == "TRUE" | cases_3393 == "TRUE")

#245157 individuals are controls for either Z947 or 3393
hearing_aid_control <- hearing_all %>% 
 filter(reclass_aid == "FALSE" & control_3393 == "FALSE")


#merge cases and controls
hearing_all <- hearing_all %>% 
  mutate(hearing_aid_cat= case_when(
    reclass_aid == "TRUE" | cases_3393 == "TRUE" ~ "case",
    reclass_aid == "FALSE" & control_3393 == "FALSE" ~ "control",
    TRUE ~"NA")
    ) 


#258965 are either cases or controls after reclassification
hear_aid <- hearing_all %>% 
  filter( hearing_aid_cat == "case" | hearing_aid_cat == "control")

```






### step 2: get the ages for hearing aids (3393)
### Extract age for Control (3393)
```{r age for control,warning=FALSE, collapse=FALSE,echo=T}
aid_age_control <- hearing_all %>% 
  filter(hearing_aid_cat == "control") 


#find out the age at the last visit (control)

offset = which(colnames(aid_age_control) == 'f.21003.0.0') - which(colnames(aid_age_control) == 'f.3393.0.0')

aid_age_control$age_aid = apply(aid_age_control, 1, function(x) {
  hear_aid = which(x[grep("f.3393", names(x))] == "No")
  first_index_offset = grep("f.3393", names(x))[1] - 1
  unlist(x[hear_aid[length(hear_aid)] + first_index_offset + offset])
})

res<-head(aid_age_control)

```


### Extract age for Cases (3393)
```{r age for cases,warning=FALSE, collapse=FALSE,echo=T}
#first category of cases (3393 true, Z974 False)
aid_age_case1 <- hearing_all %>% 
  filter(hearing_aid_cat == "case" & reclass_aid == "FALSE") 

#find out the age at the first visit (case)

offset = which(colnames(aid_age_case1) == 'f.21003.0.0') - which(colnames(aid_age_case1) == 'f.3393.0.0')

aid_age_case1$age_aid = apply(aid_age_case1, 1, function(x) {
  hear_aid =  which(x[grep("f.3393", names(x))] == "Yes")
  first_index_offset = grep("f.3393", names(x))[1] - 1
  unlist(x[min(hear_aid) + first_index_offset + offset])
})

res<-head(aid_age_case1)

#second category of cases (3393 false, Z974 true)
aid_age_case2 <- hearing_all %>% 
  filter(hearing_aid_cat == "case" & reclass_aid == "TRUE") 

#age for hearing aid based on Z974
offset = which(colnames(aid_age_case2) == 'f.41280.0.0') - which(colnames(aid_age_case2) == 'f.41270.0.0')

aid_age_case2$age_aid = apply(aid_age_case2, 1, function(x) {
  hear_aid = which(x[grep("f.41270", names(x))] == "Z974")
  first_index_offset = grep("f.41270", names(x))[1] - 1
  unlist(x[hear_aid[length(hear_aid)] + first_index_offset + offset])
})

hearing_aid_new <- aid_age_case2%>% 
  separate(age_aid, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(num_year=as.numeric(year))
  
hearing_aid_new2 <- hearing_aid_new %>%   
  mutate(age_aid= num_year - f.34.0.0) %>% 
  select(-year, -month, -day, -num_year)

#merge age for cases and controls
hearing_clean <- rbind(aid_age_case1,hearing_aid_new2, aid_age_control) 
dim(hearing_clean)

#merge back to the origical data set
hearing_clean <- merge(x = hearing_all, y = hearing_clean, by = c("IID","FID") ,all.x = TRUE) %>% 
  select(-ends_with(".y")) %>% 
  dplyr::rename_all(
    ~stringr::str_replace_all(., ".x", "")
    )
```




### Hearing difficulty/problems (2247)
### step1: classify cases and controls
```{r get cases and controls, warning=FALSE, collapse=FALSE, echo=T}

#classify cases of hearing difficulty/problems based on 2247
hearing_diff <- hearing_clean %>% 
  mutate(cases_2247 = apply(select(.,starts_with("f.2247")), 1, function(x) length(which(x == "Yes")) > 0 & max(which(x != "Yes")) < min(which(x == "Yes")))
  )

#classify controls of hearing difficulty/problems based on 2247
hearing_diff$control_2247 = with(hearing_diff, ifelse(f.2247.0.0 %in% c("No",NA) & f.2247.1.0 %in% c("No", NA) & f.2247.2.0 %in% c("No",NA) & f.2247.3.0 %in% c("No",NA) 
                                                 & !(f.2247.0.0 %in% c(NA) & f.2247.1.0 %in% c(NA) & f.2247.2.0 %in% c(NA) & f.2247.3.0 %in% c(NA)),"FALSE", "NA")) 


#380 individuals are true for Z974 but controls/NA for 2247
hearing_diff_trial <- hearing_diff %>% 
  filter(cases_Z974 == "TRUE" & (control_2247 == "FALSE" | control_2247 == "NA" ))

#reclassify based on Z974 and 2247
hearing_diff  <- hearing_diff  %>% 
  mutate(reclass_diff= case_when(
    cases_Z974 == "TRUE" & control_2247 == "FALSE" ~ "TRUE",
    cases_Z974 == "TRUE" & control_2247 == "NA"  ~ "TRUE",
    TRUE ~"FALSE") ) 


#merge cases and controls
hearing_diff <- hearing_diff %>% 
  mutate(hearing_diff_cat = case_when(
    cases_2247 == "TRUE"  ~ "case",
    control_2247 == "FALSE" & cases_Z974 == "TRUE" ~"case",
    control_2247 == "FALSE" & cases_3393 == "TRUE" ~"case",
    reclass_diff == "FALSE" & control_2247 == "FALSE" ~ "control",
    TRUE ~"NA") )

#101441 individuals are cases for either Z947 or 2247
hearing_diff_cases <- hearing_diff %>% 
 filter(reclass_diff == "TRUE" | cases_2247 == "TRUE")

#275593 individuals are controls for either Z947 or 2247
hearing_diff_control <- hearing_diff %>% 
 filter(reclass_diff == "FALSE" & control_2247 == "FALSE")

#377034 are either cases or controls after reclassification
hearing_diff_test <- hearing_diff %>% 
  filter( hearing_diff_cat == "case" | hearing_diff_cat == "control")

#check inconsistence of 2247 and hearing aid
check_inconsistence <- hearing_diff %>% 
  filter(hearing_diff_cat == "control" & hearing_aid_cat == "case")           
```


### Step 2: get the ages for Hearing difficulty/problems (2247)
### Extract age for Control (2247)
```{r age for control,warning=FALSE, collapse=FALSE,echo=T}
hearing_diff_age_control <- hearing_diff %>% 
  filter(hearing_diff_cat == "control") 

#get the age at last visit for control
offset = which(colnames(hearing_diff_age_control) == 'f.21003.0.0') - which(colnames(hearing_diff_age_control) == 'f.2247.0.0')

hearing_diff_age_control$age_diff = apply(hearing_diff_age_control, 1, function(x) {
  hear_aid = which(x[grep("f.2247", names(x))] == "No")
  first_index_offset = grep("f.2247", names(x))[1] - 1
  unlist(x[max(hear_aid) + first_index_offset + offset])
})

res<-head(hearing_diff_age_control)

```



### Extract age for Cases (2247)
```{r age for cases,warning=FALSE, collapse=FALSE,echo=T}
#case source 1: Z974 case/control, 2247 true -- based on 2247
source1 <- hearing_diff %>% 
  filter(cases_2247 == "TRUE") 

offset = which(colnames(source1) == 'f.21003.0.0') - which(colnames(source1) == 'f.2247.0.0')

source1$age_diff = apply(source1, 1, function(x) {
  hear_aid =  which(x[grep("f.2247", names(x))] == "Yes")
  first_index_offset = grep("f.2247", names(x))[1] - 1
  unlist(x[min(hear_aid) + first_index_offset + offset])
})

#case source 2: Z974 case, 2247 control -- based on Z974
source2 <- hearing_diff %>% 
  filter(control_2247 == "FALSE" & cases_Z974 == "TRUE")  

offset = which(colnames(source2) == 'f.41280.0.0') - which(colnames(source2) == 'f.41270.0.0')

source2$age_diff = apply(source2, 1, function(x) {
  hear_aid = which(x[grep("f.41270", names(x))] == "Z974")
  first_index_offset = grep("f.41270", names(x))[1] - 1
  unlist(x[hear_aid[length(hear_aid)] + first_index_offset + offset])
})

hearing_diff_new <- source2 %>% 
  separate(age_diff, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(num_year=as.numeric(year))
  
hearing_diff_new2 <- hearing_diff_new %>%   
  mutate(age_diff= num_year - f.34.0.0) %>% 
  select(-year, -month, -day, -num_year)

#case source 3: 3393 case, 2247 control -- based on 3393
source3 <- hearing_diff %>% 
  filter(control_2247 == "FALSE" & cases_3393 == "TRUE") 

offset = which(colnames(source3) == 'f.21003.0.0') - which(colnames(source3) == 'f.3393.0.0')

source3$age_diff = apply(source3, 1, function(x) {
  hear_aid =  which(x[grep("f.3393", names(x))] == "Yes")
  first_index_offset = grep("f.3393", names(x))[1] - 1
  unlist(x[min(hear_aid) + first_index_offset + offset])
})

#merge the four age datasets (cases + controls) for 2247
hearing_diff_clean <- rbind(source1, hearing_diff_age_control,source3,hearing_diff_new2) 

dim(hearing_diff_clean)

#merge the dataset (2247) back to the original dataset
hearing_diff_clean <- merge(x = hearing_clean, y = hearing_diff_clean, by = c("IID","FID") ,all.x = TRUE) %>%   
  select(-ends_with(".y")) %>% 
  dplyr::rename_all(
    ~stringr::str_replace_all(., ".x", "")
    )
```



### hearing difficulty/background noise (2257)
### step1: classify cases and controls
```{r get cases and controls, warning=FALSE, collapse=FALSE, echo=T}
hearing_noise <- hearing_diff_clean %>% 
  mutate(cases_2257 = apply(select(.,starts_with("f.2257")), 1, function(x) length(which(x == "Yes")) > 0 & max(which(x != "Yes")) < min(which(x == "Yes")))
  )

hearing_noise$control_2257 = with(hearing_noise, ifelse(f.2257.0.0 %in% c("No",NA) & f.2257.1.0 %in% c("No", NA) & f.2257.2.0 %in% c("No",NA) & f.2257.3.0 %in% c("No",NA) 
                                                 & !(f.2257.0.0 %in% c(NA) & f.2257.1.0 %in% c(NA) & f.2257.2.0 %in% c(NA) & f.2257.3.0 %in% c(NA)),"FALSE", NA)) 

#380 individuals are true for Z974 but controls/NA for 2257
hearing_noise_trial <- hearing_noise %>% 
  filter(cases_Z974 == "TRUE" & (control_2257 == "FALSE" | control_2257 == "NA" ))

#reclassify based on Z974 and 2257
hearing_noise <- hearing_noise  %>% 
  mutate(reclass_noise= case_when(
    cases_Z974 == "TRUE" & control_2257 == "FALSE" ~ "TRUE",
    cases_Z974 == "TRUE" & control_2257 == "NA"  ~ "TRUE",
    TRUE ~"FALSE") ) 


#merge cases and controls
hearing_noise <- hearing_noise %>% 
  mutate(hearing_noise_cat = case_when(
    cases_2257 == "TRUE"  ~ "case",
    control_2257 == "FALSE" & cases_Z974 == "TRUE" ~"case",
    control_2257 == "FALSE" & cases_3393 == "TRUE" ~"case",
    reclass_noise == "FALSE" & control_2257 == "FALSE" ~ "control",
    TRUE ~"NA") )

#148673 individuals are cases for either Z947 or 2257
hearing_noise_cases <- hearing_noise %>% 
 filter(reclass_noise == "TRUE" | cases_2257 == "TRUE")

#235712 individuals are controls for either Z947 or 2257
hearing_noise_control <- hearing_noise %>% 
 filter(reclass_noise == "FALSE" & control_2257 == "FALSE")

#384385 are either cases or controls after reclassification
hearing_noise_test <- hearing_noise %>% 
  filter( hearing_noise_cat == "case" | hearing_noise_cat == "control")

#check inconsistence of 2257 and hearing aid
check_inconsistence <- hearing_noise %>% 
  filter(hearing_noise_cat == "control" & hearing_aid_cat == "case")           
```


### Step 2: get the ages for Hearing difficulty/background noise (2257)
### Extract age for Control (2257)
```{r age for control,warning=FALSE, collapse=FALSE,echo=T}
hearing_noise_age_control <- hearing_noise %>% 
  filter(hearing_noise_cat == "control") 

#get the age at last visit for control
offset = which(colnames(hearing_noise_age_control) == 'f.21003.0.0') - which(colnames(hearing_noise_age_control) == 'f.2257.0.0')

hearing_noise_age_control$age_noise = apply(hearing_noise_age_control, 1, function(x) {
  hear_aid = which(x[grep("f.2257", names(x))] == "No")
  first_index_offset = grep("f.2257", names(x))[1] - 1
  unlist(x[max(hear_aid) + first_index_offset + offset])
})

res<-head(hearing_noise_age_control)

```



### Extract age for Cases (2247)
```{r age for cases,warning=FALSE, collapse=FALSE,echo=T}
#case source 1: Z974 case/control, 2257 true -- based on 2247
source1_noise <- hearing_noise %>% 
  filter(cases_2257 == "TRUE") 

offset = which(colnames(source1_noise) == 'f.21003.0.0') - which(colnames(source1_noise) == 'f.2257.0.0')

source1_noise$age_noise = apply(source1_noise, 1, function(x) {
  hear_aid =  which(x[grep("f.2257", names(x))] == "Yes")
  first_index_offset = grep("f.2257", names(x))[1] - 1
  unlist(x[min(hear_aid) + first_index_offset + offset])
})

#case source 2: Z974 case, 2257 control -- based on Z974
source2_noise <- hearing_noise %>% 
  filter(control_2257 == "FALSE" & cases_Z974 == "TRUE")  

offset = which(colnames(source2_noise) == 'f.41280.0.0') - which(colnames(source2_noise) == 'f.41270.0.0')

source2_noise$age_noise = apply(source2_noise, 1, function(x) {
  hear_aid = which(x[grep("f.41270", names(x))] == "Z974")
  first_index_offset = grep("f.41270", names(x))[1] - 1
  unlist(x[hear_aid[length(hear_aid)] + first_index_offset + offset])
})

hearing_noise_new <- source2_noise %>% 
  separate(age_noise, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(num_year=as.numeric(year))
  
hearing_noise_new2 <- hearing_noise_new %>%   
  mutate(age_noise= num_year - f.34.0.0) %>% 
  select(-year, -month, -day, -num_year)

#case source 3: 3393 case, 2257 control -- based on 3393
source3_noise <- hearing_noise %>% 
  filter(control_2257 == "FALSE" & cases_3393 == "TRUE") 

offset = which(colnames(source3_noise) == 'f.21003.0.0') - which(colnames(source3_noise) == 'f.3393.0.0')

source3_noise$age_noise = apply(source3_noise, 1, function(x) {
  hear_aid =which(x[grep("f.3393", names(x))] == "Yes")
  first_index_offset = grep("f.3393", names(x))[1] - 1
  unlist(x[min(hear_aid) + first_index_offset + offset])
})

#merge the four age datasets (cases + controls) for 2257
hearing_noise_clean <- rbind(source1_noise, hearing_noise_age_control,source3_noise,hearing_noise_new2) 

dim(hearing_diff_clean)

#merge the dataset (2257) back to the original dataset
hearing_noise_clean <- merge(x = hearing_clean, y = hearing_noise_clean, by = c("IID","FID") ,all.x = TRUE) %>%   
  select(-ends_with(".y")) %>% 
  dplyr::rename_all(
    ~stringr::str_replace_all(., ".x", "")
    )
```